---
title: "Assesment_2.2_1051525"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
options(warn=-1)
```
# Penguin submission
```{r pressure, echo=FALSE, include = FALSE}
source("scripts/import_clean.r")
source("scripts/analysis.r")
source("scripts/penguinplot1.r")
```
## Analysis of model
```{r}
#analysis
analysis1

```

### Normal Q-Q plot
Points fall roughly along the straight diagonal line, with minor tailing off at either end, however, this does not violate the assumption that the residuals are normally distributed
```{r, }
analysis2
```

### Scale-location plot
The line is roughly horizontal line with a minor positive and no major deviations;  the points are also relatively evenly spread, suggesting acceptable homogeneity with no violation of equal variance.



```{r penguinplot1, message = FALSE}
#plot graph and save to as pdf in figures
penguinplot1
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
